A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-05-21 paper/paper.tex
introduction: ongoing

A family of sets can commonly be represented in two ways: as a family indexed over a set properly speaking or as a set together with a map telling for each element of the set on which index it depends on. In the latter case, the set associated to a given index is the fiber, i.e. inverse image, of the index under the map. The two presentations are equivalent, what can be formally stated in a type-theoretic language as a fibered/indexed correspondence\footnote{At a discussion on the Zulip channel for the HoTT community, alternative names fibration/family correspondence and unbundled/bundled correspondence were proposed, among others. In our experience, the opposition fibered/indexed is spontaneously understood by many listeners. In particular, it echoes to the Grothendieck construction of fibered categories from indexed categories. Some people actually calls this correspondence ``the Grothendieck construction for dummies''. Its proof requires univalence (see~\cite{hottbook}).}:
$$(\Sigma S: \U. (S \rightarrow T)) \simeq (T \rightarrow \U)$$
where $\U$ represents the type of sets.

While a presheaf over an index category, i.e. a family of sets with maps between them according to the morphisms of the category, can be seen as living on the indexed side of the correspondence, with discrete Grothendieck fibrations on the other side of the correspondence, there are situations where the presheaf can also be seen as living on the fibered side of the correspondence. Indeed, when the index category is direct, i.e. with a downwards-well-founded collection of morphisms, as is the case of semi-simplicial and semi-cubical sets, each set serving as domain of a map can be turned into a family indexed on the codomain of the map, and, since the collection of maps is well-founded, this can be iterated, eventually producing a collection of families which we call an indexed presentation of the presheaf over the direct category.

@@

The idea for such indexed presentation of presheaves over a direct category was introduced at the Univalent Foundations year in the context of defining semi-simplicial types~\cite{}. A couple of constructions were proposed since then. One, by Voevodsky~\cite{voevodsky12}, relies on the presentation of semi-simplicial sets as a presheaf over increasing injective maps between finite ordinals. Another, by the first author, and fully formalized in Coq, relies on the presentation of semi-simplicial sets as a presheaf over face maps~\cite{herbelin15}. Another one by Part and Luo, and formalized in the Plastic proof assistant, works generically for any Reedy category with only upwards maps~\cite{part15}.

>>>>

A family of sets can commonly be represented in two ways: as a family indexed over a set properly speaking or as a set together with a map telling for each element of the set on which index it depends. In the latter case, the set associated to a given index is the fiber, i.e. inverse image, of the index under the map. The two presentations are equivalent, what can be formally stated in a type-theoretic language as a fibered/indexed correspondence\footnote{At a discussion on the Zulip channel for the HoTT community, alternative names fibration/family correspondence and unbundled/bundled correspondence were proposed, among others. In our experience, the opposition fibered/indexed is spontaneously understood by many listeners. In particular, it echoes to the Grothendieck construction of fibered categories from indexed categories. Some people actually calls this correspondence ``the Grothendieck construction for dummies''. Its proof requires univalence (see~\cite{hottbook}).}:
$$(\Sigma S: \U. (S \rightarrow T)) \simeq (T \rightarrow \U)$$
where $\U$ represents the subset of types with proof-irrelevant equality.

While a presheaf over an index category, i.e. a family of sets with maps between them according to the morphisms of the category, can be seen as living on the indexed side of the correspondence, with discrete Grothendieck fibrations on the other side of the correspondence, there are situations where the presheaf can also be seen as living on the fibered side of the correspondence. Indeed, when the index category is direct, i.e. with a downwards-well-founded collection of morphisms, as is the case of semi-simplicial and semi-cubical sets, each $F(X)$ serving as source of a map $F(X) \stackrel{F(f)}{\rightarrow} F(Y)$ can be turned into a family $F(X)$ indexed on $F(Y)$, and, since the collection of maps is well-founded, this can be iterated, eventually producing a collection of families which we call an indexed presentation of the presheaf over the direct category.

@@

The idea for such indexed presentation of presheaves over a direct category was introduced at the Univalent Foundations year in the context of defining semi-simplicial types~\cite{}. A couple of constructions were proposed since then. One, by Voevodsky~\cite{voevodsky12}, relies on the presentation of semi-simplicial sets as a presheaf over increasing injective maps between finite ordinals. Another, by the first author, and fully formalized in Coq, relies on the presentation of semi-simplicial sets as a presheaf over face maps~\cite{herbelin15}\footnote{The title of the paper ``A dependently-typed construction of semi-simplicial types'' is confusing: it was part of a project at defining semi-simplicial types, but, in practice, the definition is done in a type theory with Unicity of Identity Proofs (UIP), meaning in a type theory where types (intended to represent arbitrary $\infty$-groupoids) are actually sets ($0$-truncated $\infty$-groupoids, that is, morally, $0$-groupoids).}. Another one by Part and Luo, and formalized in the Plastic proof assistant, works generically for any Reedy category with only upwards maps~\cite{part15}.
